Left-linear rules in adhesive categories

### 2024-02-11 icalp/fusione.tex
no need of useless shifts

% direct shift
\newcommand{\shift}[1][]{\ensuremath{\mathrel{{\leftrightsquigarrow}^{\mathit{sh}}_{#1}}}}

% shift preorder
\newcommand{\shiftpre}[1][]{\ensuremath{\mathrel{{\sqsubseteq}^{\mathit{sh}}_{#1}}}}

@@

\begin{definition}[Sequential independence]
  \label{de:sequential-independence}
  Let with $\X$ an $\mathcal{M}$-adhesive category 
  and $(\X, \R)$ a left-linear DPO rewriting system. 
  Let also
  $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation of length $2$, with
  $\dder{D}_i$ using rule $\rho_i = (l_i,r_i)$.  An \emph{independence
    pair} between $\dder{D}_0$ and $\dder{D}_1$, is a pair of arrows
  $i_0\colon R_0\to D_1$ and $i_1\colon L_1\to D_0$ such that the
  following diagram commutes
  \[\xymatrix@C=15pt{L_0 \ar[d]_{m_0}&& K_0
      \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
      \ar@{.>}@/^.35cm/[drrr]|(.3)\hole_(.4){i_0}
      \ar[dr]|(.3)\hole_{h_0} && L_1 \ar@{.>}@/_.35cm/[dlll]^(.4){i_1}
      \ar[dl]|(.3)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1}
      \ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
      D_0 \ar[rr]_{g_0}&& G_1 && \ar@{>->}[ll]^{f_1} D_1
      \ar[rr]_{g_1}&& G_2}\] We will say that $\dder{D}_0$ and
  $\dder{D}_1$ are
  % \emph{weakly sequentially independent} if an independence pair
  % exists. If such independence pair is unique we will say that
  % $\dder{D}$ and $\dder{D'}$ are
  \emph{sequentially independent}.\todo{Eliminata la nozione di weakly
    indepedendent, giusto?}
\end{definition}

@@

\begin{definition}[Switch equivalence]
  \label{de:switch-equivalence}
  Let $(\X, \R)$ be a left-linear DPO rewriting system.  Let
  $\der{D}, \der{E} : G \Mapsto H$ be derivations with the same
  length, $\der{D}=\{\dder{D}_{i}\}_{i=0}^n$ and
  $\der{E}=\{\dder{E}_{i}\}_{i=0}^n$. If
  $\dder{D}_i \cdot \dder{D}_{i+1}$ is a switch of
  $\dder{E}_i \cdot \dder{E}_{i+1}$ for some $i \in [0,n-1]$ we write
  $\der{D} \shift_{(i,i+1)} \der{E}$. Then \emph{switch
    equivalence} is the least equivalence on derivation such that if
  $\der{D} \leftrightsquigarrow_{(i,i+1)} \der{E}$ then
  $\der{D} \equiv^s \der{E}$.
\end{definition}

@@

\subsection{The $3$-steps lemmas}

In the case of linear rewriting systems, the characterisation of switch equivalence based on consistent
permutation allows to derive some simple but
fundamental results for the theory, namely
\begin{itemize}
\item the result of switching does not depends on the order, but only
  on the permutation;
\item independence is global, meaning that if two independent steps are
  moved forward or backward in a derivation, they remain
  independent.
\end{itemize}

The failure of the characterisation for left-linear rules poses the
question asking whether the results above also fail.
%
We observe that
\begin{itemize}

\item the first can be proved, yet a different proof technique is needed,
  taking advantage from the axiomatic characterisation of switching;

\item Independence is global partially fails: one part holds, the
  other in a weak form, but this is conceptually fini (related to or
  causality).
\end{itemize}

As a beginning step in our analysis of switch equivalence, we will
establish three lemmas dealing with derivations of length
$3$.

\begin{restatable}[First $3$-steps Lemma]{lemma}{lemThreeSteps}
  \label{lem:three-steps}
  Let $\X$ be an $\mathcal{M}$-adhesive category and
  $(\X,\R)$ a left-linear DPO rewriting system.
%
  Consider a
  derivation $\der{D}=\{\dder{D}_i\}_{i=0}^2$ and suppose that
  $(i_0,i_1)$ is a good independence pair between $\dder{D}_0$ and
  $\dder{D}_1$, $(a_0,a_1)$ one between $\dder{D}_1$ and $\dder{D}_2$
  and $(e_0, e_1)$ one between $\dder{D}_0$ and
  $S_{a_0,a_1}(\dder{D}_2)$.
\end{restatable}

[Proof~\ref{lemThreeSteps-proof}]

\rem{remove}{
\begin{lemma}[Second $3$-steps Lemma]\label{lem:secondo}
	contenuto...
\end{lemma}
\begin{proof}
	contenuto...
\end{proof}

\begin{lemma}[Third $3$-steps Lemma]\label{lem:terzo}
	contenuto...
\end{lemma}
\begin{proof}
	contenuto...
\end{proof}
}

\todo{dividere in due il lemma dopo. devono venire tre lemmi}
\begin{restatable}[Second 3-steps Lemma]{lemma}{lemIndepGlobalLeft}
  \label{lem:indep-global-left}
  Let $\X$ be an $\mathcal{M}$-adhesive category and
  $(\X,\R)$ a left-linear DPO rewriting system.
  %
  Consider a derivation
  $\der{D}=\{\dder{D}_i\}_{i=0}^2$ and suppose that $(i_0,i_1)$ is a
  good independence pair between $\dder{D}_0$ and $\dder{D}_1$,
  $(a_0,a_1)$ one between $\dder{D}_1$ and $\dder{D}_2$ and
  $(e_0, e_1)$ one between $\dder{D}_0$ and
  $S_{a_0,a_1}(\dder{D}_2)$. Then the following properties hold true.
  \begin{enumerate}
  \item $S_{e_0,e_1}(\dder{D}_0)$ and $S_{a_0,a_1}(\dder{D}_1)$ are
     sequentially independent.
  \item If $S_{i_0, i_1}(\dder{D}_0)\updownarrow_! \dder{D}_2$ with a
    good independence pair $(\alpha_0, \alpha_1)$, then
    $S_{i_0,i_1}(\dder{D}_1)$ and $S_{\alpha_0, \alpha_1}(\dder{D}_2)$
    are  sequentially independent.
  \end{enumerate}
\end{restatable}

@@

\lemThreeSteps*
\label{lemThreeSteps-proof}

@@

>>>>

% sequential independence
\newcommand{\seqind}{\ensuremath{\updownarrow}}

% direct shift
\newcommand{\shift}[1]{\ensuremath{\mathrel{{\leftrightsquigarrow}_{#1}}}}

@@

\begin{definition}[Sequential independence]
  \label{de:sequential-independence}
  Let with $\X$ an $\mathcal{M}$-adhesive category 
  and $(\X, \R)$ a left-linear DPO rewriting system. 
  Let also
  $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation of length $2$, with
  $\dder{D}_i$ using rule $\rho_i = (l_i,r_i)$.  An \emph{independence
    pair} between $\dder{D}_0$ and $\dder{D}_1$, is a pair of arrows
  $i_0\colon R_0\to D_1$ and $i_1\colon L_1\to D_0$ such that the
  following diagram commutes
  \[\xymatrix@C=15pt{L_0 \ar[d]_{m_0}&& K_0
      \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
      \ar@{.>}@/^.35cm/[drrr]|(.3)\hole_(.4){i_0}
      \ar[dr]|(.3)\hole_{h_0} && L_1 \ar@{.>}@/_.35cm/[dlll]^(.4){i_1}
      \ar[dl]|(.3)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1}
      \ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
      D_0 \ar[rr]_{g_0}&& G_1 && \ar@{>->}[ll]^{f_1} D_1
      \ar[rr]_{g_1}&& G_2}\] We will say that $\dder{D}_0$ and
  $\dder{D}_1$ are
  % \emph{weakly sequentially independent} if an independence pair
  % exists. If such independence pair is unique we will say that
  % $\dder{D}$ and $\dder{D'}$ are
  \emph{sequentially independent} and write $\dder{D}_0 \seqind \dder{D}_1$.\todo{Eliminata la nozione di weakly
    indepedendent, giusto?}
\end{definition}

@@

\begin{definition}[Switch equivalence]
  \label{de:switch-equivalence}
  Let $(\X, \R)$ be a left-linear DPO rewriting system.  Let
  $\der{D}, \der{E} : G \Mapsto H$ be derivations with the same
  length, $\der{D}=\{\dder{D}_{i}\}_{i=0}^n$ and
  $\der{E}=\{\dder{E}_{i}\}_{i=0}^n$. If
  $\dder{D}_i \cdot \dder{D}_{i+1}$ is a switch of
  $\dder{E}_i \cdot \dder{E}_{i+1}$ for some $i \in [0,n-1]$ we write
  $\der{D} \shift{\transp{i}} \der{E}$. Then \emph{switch
    equivalence} is the least equivalence on derivation such that if
  $\der{D} \shift{\transp{i}} \der{E}$ then
  $\der{D} \equiv^s \der{E}$.
\end{definition}

@@

\subsection{A Canonical Form for  Switch Equivalences}

In the case of linear rewriting systems, the characterisation of
switch equivalence based on consistent permutations is the key to
derive some simple but fundamental properties for the theory, namely
\begin{enumerate}
\item \emph{Confluence of switching}: when transforming derivation
  sequences into equivalent ones by switching independent steps, the
  result not depends on the order in which switchings are applied, but
  only on the resulting permutation;

\item \emph{Globality of independence}: if two independent steps are
  moved forward or backward due to the application of switchings to a
  derivation, they remain independent.
\end{enumerate}

In particular, the two properties allows one to obtain a \emph{canonical form} for switching sequences: whenever two derivation sequences are switch equivalence via some permutation $\sigma$, then each one can be obtained from the other by applying switchings to pairs of rule applications which $\sigma$ reverses (the so-called inversions).\todo{P: questa prima parte da riportare in parte in intro}

The failure of the characterisation of switch equivalence via
consistent permutations for left-linear rules naturally leads to the
question asking whether also the properties above cease to hold.

In this section we show that actually much of the theory can be retained: confluence of switching still holds, globality of independence holds in a weak form conceptually motivated by the fact that fusions introduce disjubctive causality. Finally a canonical form for switching sequences can also be recovered.

We start by proving two lemmas dealing with derivations of length
$3$. Despite being technical, these lemma are the fundamental building blocks for showing properties (1) and (2) above.

The first lemma capture the essence of the confluence of
switchings. It shows that if we have a derivation consisting of three
rewriting steps, say $r_0\, r_1\, r_2$ and we switch them leading to the
sequence $r_2\, r_1\, r_0$, we obtain the same resulting derivation,
indepdendently from the order of the switchings.

\begin{restatable}[confluence of switchings]{lemma}{lemSwitchConfluence}
  \label{lem:switch-confluence}
  Let $\X$ be an $\mathcal{M}$-adhesive category and $(\X,\R)$ a
  left-linear DPO rewriting system.
  % 
  Consider a derivation $\der{D}=\{\dder{D}_i\}_{i=0}^2$ and suppose
  that $(i_0,i_1)$ is a good independence pair between $\dder{D}_0$
  and $\dder{D}_1$, $(a_0,a_1)$ one between $\dder{D}_1$ and
  $\dder{D}_2$ and $(e_0, e_1)$ one between $\dder{D}_0$ and
  $S_{a_0,a_1}(\dder{D}_2)$.\todo{P: non e' meglio qui limitarci a very tame, cosi' da evitare riferimenti a good pairs?}
\end{restatable}

[Proof~\ref{lemSwitchConfluence-proof}]

The next technical lemma is at the core of globality of independence. It shows that if we have a derivation consisting of three
rewriting steps, say $r_0\, r_1\, r_2$ where $r_0$ and $r_1$ are independent, then if we switch $r_2$ to the first position, obtaining a derivation $r_2\, r_0\, r_1$, the steps $r_0$ and $r_1$ are still independent.

Instead, if in $r_0\, r_1\, r_2$ we have $r_1$ and $r_2$ are
independent, and we switch $r_0$ to the last position, obtaining a
derivation $r_1\, r_2\, r_0$, the steps $r_1$ and $r_2$ could be no
longer independent, unless an additional condition is
satisfied. Intuitively this is due to the fact that, dealing with
left-linear rewriting systems introduces disjunctive forms of
causality. Hence, if in $r_0\, r_1\, r_2$ we have that $r_0, r_1$ are
disjunctive causes of $r_2$, at least one of the two is needed to
enable the application of $r_2$ and this explain why in
$r_1\, r_2\, r_0$ we have that $r_1$ and $r_2$ are no longer
independent. \todo{P: da dire meglio e Ci vorrebbe un esempio}

\begin{restatable}[Globality of indepedence]{lemma}{lemIndepGlobalLeft}
  \label{lem:indep-global-left}
  Let $\X$ be an $\mathcal{M}$-adhesive category and
  $(\X,\R)$ a left-linear DPO rewriting system.
  Consider a derivation
  $\der{D}=\{\dder{D}_i\}_{i=0}^2$ and suppose that $(i_0,i_1)$ is a
  good independence pair between $\dder{D}_0$ and $\dder{D}_1$,
  $(a_0,a_1)$ one between $\dder{D}_1$ and $\dder{D}_2$ and
  $(e_0, e_1)$ one between $\dder{D}_0$ and
  $S_{a_0,a_1}(\dder{D}_2)$. Then the following properties hold true.
  \begin{enumerate}
  \item
    \label{lem:indep-global-left:1}
    $S_{e_0,e_1}(\dder{D}_0)$ and $S_{a_0,a_1}(\dder{D}_1)$ are
    sequentially independent.

  \item
    \label{lem:indep-global-left:2}
    If $S_{i_0, i_1}(\dder{D}_0)\seqind \dder{D}_2$ with a
    good independence pair $(\alpha_0, \alpha_1)$, then
    $S_{i_0,i_1}(\dder{D}_1)$ and $S_{\alpha_0, \alpha_1}(\dder{D}_2)$
    are  sequentially independent.
  \end{enumerate}
\end{restatable}

\todo{P: anche qui sopra mi limiterei a very tame, cos\i' da evitare riferimenti a good e se si riesce anche alle indepdencen pairs}
  %

@@

\lemSwitchConfluence*
\label{lemSwitchConfluence-proof}

@@

%%%%%%%%%%% NO NEED OF USELESS SHIFTS

We next prove some technical lemma which will be useful to show the
possibility of transforming a derivation into a switch equivalent one
by switching only step along inversions.

\begin{lemma}
  \label{le:switchA}
  Let $\der{D}_i$ for $i \in \interval[0]{n}$ be derivation sequences
  and consider a switching sequence consisting only of inversions
  \begin{center}
    $\der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n$
  \end{center}
  Let $i \in \interval{n}$ be such that
  $\transp{i} \in \inv{\nu_{1,n}}$ and for all $k > i+1$ it holds
  $\nu_{1,n}(k) > \nu_{1,n}(i+1)$.  Then there is a switching sequence
  consisting only of inversions which starts with $\transp{i}$, i.e.,
  \begin{center}
    $\der{D}_0 \shift{\transp{i}} \der{D}_1' \shift{\nu_2'} \ldots
    \shift{\nu_n'} \der{D}_n$
  \end{center}
\end{lemma}

\begin{proof}
  We proceed by induction on $n$. If $n=1$, necessarily
  $\nu_1 =  \transp{i}$ and thus we are done.

  If $n>1$, let $\nu_1 = \transp{j}$.  We distinguish various case
  depending on how $\nu_1$ relates to $\transp{i}$.

  \bigskip
  \noindent
  (i) $j=i$\\
  %
  In this case there is nothing to prove: $\nu_1 = \transp{i}$, hence
  the original switching sequence is already of the desired shape.

  \bigskip
  \noindent
  (ii) $j<i-1$ or $j>i-1$\\
  %
  In this case $\nu_1$ does not ``interfere'' with $\transp{i}$, i.e.,
  $i$ still satisfies the hypotheses of the lemma with respect to the
  switching sequence
  $\der{D}_1 \shift{\nu_2} \ldots \shift{\nu_n} \der{D}_n$.

  Hence we can apply the inductive hypothesis and derive the existence
  of a switching sequence
  \begin{center}
    $\der{D}_1 \shift{\transp{i}} \der{D}_2' \shift{\nu_3'} \ldots \shift{\nu_n'} \der{D}_n$
  \end{center}

  Recall that $\der{D}_0 \shift{\nu_1} \der{D}_1$ and, given the
  working assumption (ii), the switches $\nu_1$ and $\transp{i}$ can
  be done in reverse order, thus getting, as desired
  \begin{center}
    $\der{D}_0 \shift{\transp{i}} \der{D}_1' \shift{\nu_1} \ldots
    \shift{\nu_n'} \der{D}_n$
  \end{center}

  \bigskip
  \noindent
  (iii) $j=i-1$, i.e., $\nu_1 = \transp{i-1}[i]$\\
  %
  In this case $\der{D}_1$ is obtained from $\der{D}_0$ by switching
  steps $i-1$ and $i$, as pictorially reported in the first two
  columns of Fig.~\ref{fi:switchA}. Since we are dealing with
  switching sequence consisting only of inversions and by hypothesis
  $\nu_{1,n}(i) > \nu_{1,n}(i+1)$, whence
  $\nu_{2,n}(i-1) > \nu_{2,n}(i+1)$, we deduce that necessarily
  $\nu_{2,n}(i) > \nu_{2,n}(i+1)$ and $\nu_{2,n}(k) > \nu_{2,n}(i+1)$
  for all $k > i+1$. Hence we can use the inductive hypothesis to
  deduce that the switching sequence from $\der{D}_1$ can start by
  switching $\transp{i}$, i.e., we obtain (see the third column in
  Fig.~\ref{fi:switchA})
  \begin{center}
    $\der{D}_1 \shift{\transp{i}} \der{D}_2' \shift{\nu_3'} \ldots
    \shift{\nu_n'} \der{D}_n$
  \end{center}

  Now, if we focus on $\der{D}_2'$ we observe that
  $\nu_{3,n}'(i-1) > \nu_{3,n}'(i)$ and
  $\nu_{3,n}'(k) > \nu_{3,n}'(i)$ for all $k > i$. Hence we can apply
  again the inductive hypothesis and get a switching sequence
    \begin{center}
    $\der{D}_2' \shift{\transp{i-1}[i]} \der{D}_3'' \shift{\nu_3''} \ldots
    \shift{\nu_n''} \der{D}_n$
  \end{center}
  as depicted in Fig.~~\ref{fi:switchA}, starting from the fourth column.

  By Lemma~\ref{lem:indep-global-left}(\ref{lem:indep-global-left:1})\todo{Check numbering} we can
  reorganize the first three switchings as

  \begin{center}
    $\der{D}_0 \shift{\transp{i}} \der{D}_1'''
    \shift{\transp{i-1}[i]} \der{D}_2'''
    \shift{\transp{i}} \der{D}_3''$
  \end{center}

  The fact that the three switchings above lead to $\der{D}_3''$ is a
  consequence of Lemma~\ref{lem:switch-confluence}.

  \bigskip

  Observe that it cannot be $j=i+1$, otherwise we would have
  $\nu_{1,n}(j+1) < \nu_{1,n}(i)$, contradicting the hypotheses. Thus
  this case does not need to be considered and the proof is completed.  
\end{proof}

\begin{figure}
  \centering
  \begin{tikzpicture}[font=\small]
    % Define arrow colors
    \def\colorA{green!60!black}
    \def\colorB{blue}
    \def\colorC{red!100}

    % Define the length of each line
    \def\vlen{5mm}
    \def\hlen{18mm}
    \def\gap{1mm}

    % First line D0 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    % Draw label on top of the line
    \node[above] at (0*\hlen,0) {$\der{D}_0$};
    \node[above] at (0.5*\hlen,0) {$\shift{\nu_1}$};

    % Draw the black part of the line
    \draw (0*\hlen,0) -- (0*\hlen,-2*\vlen+\gap);

    % Draw the first colored labeled arrow
    \draw[->, \colorA] (0*\hlen,-2*\vlen) -- (0*\hlen,-3*\vlen+\gap) node[midway, left] {$i-1$};
    % Draw the second colored labeled arrow
    \draw[->, \colorB] (0*\hlen,-3*\vlen) -- (0*\hlen,-4*\vlen+\gap) node[midway, left] {$i$};
    % Draw the third colored labeled arrow
    \draw[->, \colorC] (0*\hlen,-4*\vlen) -- (0*\hlen,-5*\vlen+\gap) node[midway, left] {$i+1$};
    % Draw the last black part of the line
    \draw (0*\hlen,-5*\vlen) -- (0*\hlen,-7*\vlen);

    % Second line D1 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    % Draw label on top of the line
    \node[above] at (1*\hlen,0) {$\der{D}_1$};
    \node[above] at (1.5*\hlen,0) {$\shift{\transp{i}}$};

    % Draw the black part of the line
    \draw (1*\hlen,0) -- (1*\hlen,-2*\vlen+\gap);

    % Draw the first colored labeled arrow
    \draw[->, \colorB] (1*\hlen,-2*\vlen) -- (1*\hlen,-3*\vlen+\gap) node[midway, left] {};
    % Draw the second colored labeled arrow
    \draw[->, \colorA] (1*\hlen,-3*\vlen) -- (1*\hlen,-4*\vlen+\gap) node[midway, left] {};
    % Draw the third colored labeled arrow
    \draw[->, \colorC] (1*\hlen,-4*\vlen) -- (1*\hlen,-5*\vlen+\gap) node[midway, left] {};
    % Draw the last black part of the line
    \draw (1*\hlen,-5*\vlen) -- (1*\hlen,-7*\vlen);

    % Repeat similar patterns for the third and fourth lines

    % Third line: d2 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \node[above] at (2*\hlen,0) {$\der{D}_2'$};
    \node[above] at (2.5*\hlen,0) {$\shift{\transp{i-1}[i]}$};

    \draw (2*\hlen,0) -- (2*\hlen,-2*\vlen+\gap);

    \draw[->, \colorB] (2*\hlen,-2*\vlen) -- (2*\hlen,-3*\vlen+\gap) node[midway, left] {};
    \draw[->, \colorC] (2*\hlen,-3*\vlen) -- (2*\hlen,-4*\vlen+\gap) node[midway, left] {};
    \draw[->, \colorA] (2*\hlen,-4*\vlen) -- (2*\hlen,-5*\vlen+\gap) node[midway, left] {};
    \draw (2*\hlen,-5*\vlen) -- (2*\hlen,-7*\vlen);

    % Fourth line %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \node[above] at (3*\hlen,0) {$\der{D}_3''$};
    \node[above] at (3.5*\hlen,0) {$\shift{\nu_3''}$};

    \draw (3*\hlen,0) -- (3*\hlen,-2*\vlen+\gap);
    \draw[->, \colorC] (3*\hlen,-2*\vlen) -- (3*\hlen,-3*\vlen+\gap) node[midway, left] {};
    \draw[->, \colorB] (3*\hlen,-3*\vlen) -- (3*\hlen,-4*\vlen+\gap) node[midway, left] {};
    \draw[->, \colorA] (3*\hlen,-4*\vlen) -- (3*\hlen,-5*\vlen+\gap) node[midway, left] {};
    \draw (3*\hlen,-5*\vlen) -- (3*\hlen,-7*\vlen);

    % Dots after the last line
    \draw (4*\hlen,0*\vlen) node[above] {$\dots$};

    % Last line %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \node[above] at (5*\hlen,0) {$\der{D}_n$};
    \node[above] at (4.5*\hlen,0) {$\shift{\nu_{n-1}}$};

  \end{tikzpicture}
  \caption{Picture for the proof of Lemma~\ref{le:switchingsA}}
  \label{fi:switchA}
\end{figure}

\begin{remark}
  \label{rem:switchA}
  Observe that given a switching sequence as in the lemma above
  $\der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
  \shift{\nu_n} \der{D}_n$, if we consider the inversion with largest index:
  \begin{center}
    $k = \max \{j \mid \transp{j} \in \inv{\nu_{1,n}}\}$
  \end{center}
  then it satisfies the hypotheses of the lemma.
\end{remark}

\begin{lemma}
  \label{le:switchB}
  Let $\der{D}_i$ for $i \in \interval[0]{n}$ be derivation sequences
  and consider a switching sequence consisting only of inversions
  \begin{center}
    $\der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n$
  \end{center}
  Let $i \in \interval{n}$ be such that
  $\transp{i} \in \inv{\nu_{1,n}}$ and assume that steps $i$ and $i+1$ in $\der{D}_0$ are sequential independent. Then there is a switching sequence
  consisting only of inversions which starts with $\transp{i}$, i.e.,
  \begin{center}
    $\der{D}_0 \shift{\transp{i}} \der{D}_1' \shift{\nu_2'} \ldots
    \shift{\nu_n'} \der{D}_n$
  \end{center}
\end{lemma}

\begin{proof}
  We proceed by induction on $n$. If $n=1$, necessarily
  $\nu_1 =  \transp{i}$ and thus we are done.

  If $n>1$, take $k = \max \{j \mid \transp{j} \in \inv{\nu_{1,n}}\}$, which,
  as observed in Remark~\ref{rem:switchA}, satisfies the hypotheses of
  Lemma~\ref{le:switchA}. Hence, by the mentioned lemma, we can start
  the switching sequence with the transposition $\transp{k}$, thus
  obtaining:
  \begin{center}
    $\der{D}_0 \shift{\transp{k}} \der{D}_1' \shift{\nu_2'} \ldots
    \shift{\nu_n'} \der{D}_n$
  \end{center}

  Moreover, since $\transp{i}$ is an inversion, we have that
  $\i \leq k$.  We distinguish various case depending on how $k$
  relates to $i$.

  \bigskip
  \noindent
  (i) $j=i$\\
  %
  In this case we are done: $\nu_1' = \transp{i}$, hence we obtained a
  switching sequence of the desired shape.

  \bigskip
  \noindent
  (ii) $k > i+1$\\
  %
  In this case the switching $\transp{k}$ does not ``interfere'' with
  $\transp{i}$, , as depicted in Fig.~\ref{fi:switchB1} (first two
  columns).

  Hence if we focus on
  \begin{center}
    $\der{D}_1' \shift{\nu_2'} \ldots
    \shift{\nu_n'} \der{D}_n$
  \end{center}
  we can observe that
  $\nu_{2,n}'(i)=\nu_{1,n}(i) > \nu_{1,n}(i+1)=\nu_{2,n}'(i+1)$.
  %
  Thus we can use the inductive hypothesis and obtain (see
  Fig.~\ref{fi:switchB1}, third column):
  \begin{center}
    $\der{D}_1' \shift{\transp{i}} \der{D}_2'' \shift{\nu_3''} \ldots
    \shift{\nu_n''} \der{D}_n$
  \end{center}

\begin{figure}
  \centering
  \begin{tikzpicture}[font=\small]
    % Define arrow colors
    \def\colorA{red!100}
    \def\colorB{blue}
    \def\colorC{green!60!black}
    \def\colorD{orange!60!black}

    % Define the length of each line
    \def\vlen{5mm}
    \def\hlen{20mm}
    \def\gap{1mm}

    % First line D0 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    % Draw label on top of the line
    \node[above] at (0*\hlen,0) {$\der{D}_0$};
    \node[above] at (0.5*\hlen,0) {$\shift{\transp{k}}$};

    % Draw the black part of the line
    \draw (0*\hlen,0) -- (0*\hlen,-1*\vlen+\gap);

    % Draw the first colored labeled arrow
    \draw[->, \colorA] (0*\hlen,-1*\vlen) -- (0*\hlen,-2*\vlen+\gap) node[midway, left] {$i$};
    % Draw the second colored labeled arrow
    \draw[->, \colorB] (0*\hlen,-2*\vlen) -- (0*\hlen,-3*\vlen+\gap) node[midway, left] {$i+1$};
    % Draw the second black part of the line
    \draw (0*\hlen,-3*\vlen) -- (0*\hlen,-4*\vlen+\gap);

    % Draw the third colored labeled arrow
    \draw[->, \colorC] (0*\hlen,-4*\vlen) -- (0*\hlen,-5*\vlen+\gap) node[midway, left] {$k$};
    % Draw the fourth colored labeled arrow
    \draw[->, \colorD] (0*\hlen,-5*\vlen) -- (0*\hlen,-6*\vlen+\gap) node[midway, left] {$k+1$};
    % Draw the last black part of the line
    \draw (0*\hlen,-6*\vlen) -- (0*\hlen,-7*\vlen);

    % Second line D1 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    % Draw label on top of the line
    \node[above] at (1*\hlen,0) {$\der{D}_1'$};
    \node[above] at (1.5*\hlen,0) {$\shift{\nu_2'}$};

    % Draw the black part of the line
    \draw (1*\hlen,0) -- (1*\hlen,-1*\vlen+\gap);

    % Draw the first colored labeled arrow
    \draw[->, \colorA] (1*\hlen,-1*\vlen) -- (1*\hlen,-2*\vlen+\gap) node[midway, left] {};
    % Draw the second colored labeled arrow
    \draw[->, \colorB] (1*\hlen,-2*\vlen) -- (1*\hlen,-3*\vlen+\gap) node[midway, left] {};
    % Draw the second black part of the line
    \draw (1*\hlen,-3*\vlen) -- (1*\hlen,-4*\vlen+\gap);

    % Draw the third colored labeled arrow
    \draw[->, \colorD] (1*\hlen,-4*\vlen) -- (1*\hlen,-5*\vlen+\gap) node[midway, left] {};
    % Draw the fourth colored labeled arrow
    \draw[->, \colorC] (1*\hlen,-5*\vlen) -- (1*\hlen,-6*\vlen+\gap) node[midway, left] {};
    % Draw the last black part of the line
    \draw (1*\hlen,-6*\vlen) -- (1*\hlen,-7*\vlen);

    % Third line D2 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    % Draw label on top of the line
    \node[above] at (2*\hlen,0) {$\der{D}_2''$};
    \node[above] at (2.5*\hlen,0) {$\shift{\nu_3''}$};

    % Draw the black part of the line
    \draw (2*\hlen,0) -- (2*\hlen,-1*\vlen+\gap);

    % Draw the first colored labeled arrow
    \draw[->, \colorB] (2*\hlen,-1*\vlen) -- (2*\hlen,-2*\vlen+\gap) node[midway, left] {};
    % Draw the second colored labeled arrow
    \draw[->, \colorA] (2*\hlen,-2*\vlen) -- (2*\hlen,-3*\vlen+\gap) node[midway, left] {};
    % Draw the second black part of the line
    \draw (2*\hlen,-3*\vlen) -- (2*\hlen,-4*\vlen+\gap);

    % Draw the third colored labeled arrow
    \draw[->, \colorC] (2*\hlen,-4*\vlen) -- (2*\hlen,-5*\vlen+\gap) node[midway, left] {};
    % Draw the fourth colored labeled arrow
    \draw[->, \colorD] (2*\hlen,-5*\vlen) -- (2*\hlen,-6*\vlen+\gap) node[midway, left] {};
    % Draw the last black part of the line
    \draw (2*\hlen,-6*\vlen) -- (2*\hlen,-7*\vlen);

    % Dots after the last line
    \draw (3*\hlen,0*\vlen) node[above] {$\dots$};

    % Last line %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \node[above] at (4*\hlen,0) {$\der{D}_n$};
    \node[above] at (3.5*\hlen,0) {$\shift{\nu_{n-1}''}$};

  \end{tikzpicture}
  \caption{Picture for the proof of Lemma~\ref{le:switchB}, item (ii)}
  \label{fi:switchB1}
\end{figure}

  Now, since $k>i+1$, the switchings $\transp{i}$ and $\transp{k}$ can
  be performed in reverse order and thus we obtain
  \begin{center}
    $\der{D}_0 \shift{\transp{i}} \der{D}_1'' \shift{\transp{k}}
    \der{D}_2'' \shift{\nu_3''} \ldots \shift{\nu_n''} \der{D}_n$
  \end{center}
  as desired.

  \bigskip
  \noindent
  (iii) $k = i+1$\\
  %
  The situation is depicted in Fig.~\ref{fi:switchB2}. In
  $\der{D}_1'$, by the choice of $k$ and the fact that all $\nu_i$ are
  inversions, one can deduce that $i$ satisfies the hyp[otheses of
  Lemma~\ref{le:switchA}. Hence we obatain (see Fig.~\ref{fi:switchB2}, first two coulmns):
  \begin{center}
    $\der{D}_1' \shift{\transp{i}} \der{D}_2'' \shift{\nu_2''}
    \der{D}_3'' \shift{\nu_4''} \ldots \shift{\nu_n''} \der{D}_n$
  \end{center}

  Since in $\der{D}_0$ steps in positions $i$ and $i+1$ are sequential
  independent, we can use
  Lemma~\ref{lem:indep-global-left}(\ref{lem:indep-global-left:1})\todo{Check
    numbering}, we have that the corresponding steps in positions
  $i+1$ and $i+2$ in $\der{D}_2''$ are stil sequential
  independent. Hence, we can use the inductive hypothesis on
  $\der{D}_2''$ and deduce the existence of a switching sequence where
  $\transp{i+1}[i+2]$ can be switched first, i.e., we obtain (see
  Fig.~\ref{fi:switchB2}, third coulmn):
  \begin{center}
    $\der{D}_2'' \shift{\transp{i+1}[i+2]}
    \der{D}_3''' \shift{\nu_4'''} \ldots \shift{\nu_n'''} \der{D}_n$
  \end{center}
